---
title: "Untitled"
author: "Ken Mwai"
date: "Thursday, June 04, 2015"
output: beamer_presentation
---

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```


## Slide with a model

```{r, echo=T}
Bloodloss<-read.csv("Bloodloss.csv", header=T)
model1<-lm(Bloodloss$bloodloss ~ Bloodloss$worm)
sumry <- summary.lm(model1)
sumry
```

## Slide using model outputs

The r squared value is `r sumry$r.squared` 


## Slide with the formula

$$
P(NT|S x) = \frac{P(NT|S) P(NT)}{P(S)} = \frac{P(S|NT) P(NT)}{P(S|NT) P(NT) + P(S|ET) P(ET)}
$$

$$
P(NT|S x) = \frac{0.9*05}{(0.9*0.5)+(0.7*0.5)} = \frac{0.45}{0.8)} = `r 0.45/0.8`
$$


## Slide with Table 


```{r, echo=F}
library(markdown)
library('knitr')
table1 <- head(Bloodloss)
```


```{r echo=F,results='asis',error=F,warning=F}
kable(table1, format = "markdown")
```
